---
title: "Data Science Midterm"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuious.color = "viridis",
  ggplot2.continuious.fill = "viridis"
  
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

The raw data used in this analysis consists of Change of Address (COA) data from the US Postal Service for New York City between 2018 and 2022. The COA data includes information on the total number of permanent address changes going into and out of each ZIP code in New York City. The initial section of the report focuses on importing, cleaning, and ensuring data quality. This process results in a clean and structured dataset, ready for analysis. The dataset's characteristics, including observations, unique ZIP codes, and neighborhoods, are described. Data quality issues are also addressed. The subsequent section visualizes the data with an analysis of address change trends, uncovering extreme values, and plots that highlight neighborhood-level insights over the five-year period. The report concludes with an acknowledgment of any limitations of the dataset.


```{r}

```

