---
title: "Data Science Midterm"
author: "Victoria Mello (vsm2118)"
date: "October 23, 2023"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

devtools::install_github("benmarwick/wordcountaddin",  type = "source", dependencies = TRUE)
library(koRpus.lang.en)
wordcountaddin::word_count()

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuious.color = "viridis",
  ggplot2.continuious.fill = "viridis"
  
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

The raw data used in this analysis consists of Change of Address (COA) data from the US Postal Service for New York City between 2018 and 2022. The COA data includes information on the total number of permanent address changes going into and out of each ZIP code in New York City. The initial section of the report focuses on importing, cleaning, and ensuring data quality for analysis. The subsequent section visualizes the data with an analysis of net address change trends across boroughs for the 5 year period and addresses limitations of the dataset.


## Section 1 – Data import, cleaning, and quality control

```{r}
zip_data = read_csv("zip_codes.csv")

sheet_names <- c("2018", "2019", "2020", "2021", "2022")
coa_data <- map_dfr(sheet_names, ~ readxl::read_xlsx("USPS.xlsx", sheet = .x))

view(coa_data)

coa_data = coa_data %>% 
  janitor::clean_names() %>% 
  mutate(
    year = as.integer(str_sub(month, 1, 4)),
    net_change = `total_perm_in` - `total_perm_out`
  )
```

```{r}
# Create a "borough" variable in the ZIP code data
zip_data <- zip_data %>%
  janitor::clean_names() %>% 
  rename(zipcode = zip_code) %>% 
  mutate(
    borough = case_when(
      county_name %in% c("Kings") ~ "Brooklyn",
      county_name %in% c("Queens") ~ "Queens",
      county_name == "Bronx" ~ "Bronx",
      county_name == "New York" ~ "Manhattan",
      county_name == "Richmond" ~ "Staten Island"
    )
  )
  view(zip_data)

```

```{r}
# Merge the COA and ZIP code data based on ZIP code
merged_data <- coa_data %>%
  left_join(zip_data, by = "zipcode")

# Select only the necessary variables for later visualizations
final_data <- merged_data %>%
    mutate(
    borough = case_when(
      county_name == "Kings" ~ "Brooklyn",
      county_name == "Queens" ~ "Queens",
      county_name == "Bronx" ~ "Bronx",
      county_name == "New York" ~ "Manhattan",
      county_name == "Richmond" ~ "Staten Island"
    ),
    month = month(month, label = TRUE)
  ) %>%
  select(zipcode, neighborhood, borough, year, month, net_change, city)

# Cleaning zipcodes incorrectly classified for borough
final_data <- final_data %>%
  mutate(borough = case_when(
    zipcode == 10463 ~ "Bronx",
    zipcode == 11201 ~ "Brooklyn",
    TRUE ~ borough
  ))
view(final_data)
```

Major steps in data wrangling involved importing, cleaning, and merging datasets. This process included transforming the Change of Address (COA) data, creating a "borough" variable for zipcode data, merging the datasets by zipcode, and selecting key variables. Data quality issues were addressed by reclassifying zipcodes incorrectly classified for borough through manual correction. The resulting tidy dataset contains a total of `r nrow(final_data)` observations. It includes data from `r n_distinct(final_data$zipcode)` zipcodes and data from `r n_distinct(final_data$neighborhood)` neighborhoods. 

```{r}
# Filter the data for Manhattan and Queens
manhattan_data <- final_data %>%
  filter(borough == "Manhattan")

queens_data <- final_data %>%
  filter(borough == "Queens")

# Create tables for the most common cities in Manhattan and Queens
manhattan_table <- manhattan_data %>%
  count(city, sort = TRUE)

queens_table <- queens_data %>%
  count(city, sort = TRUE)

knitr::kable(manhattan_table, caption = "Manhattan most common city value")
knitr::kable(queens_table, caption = "Queens most common city value")
```

#### Data Quality Issues 
In the Manhattan table, variations in city names like "NEW YORK" and "NYC" can lead to inaccurate analysis and visualization, potentially duplicating or misrepresenting zipcode data. "BRONX" and "BROOKLYN" in the Manhattan table indicate data entry errors or misclassification, affecting borough-specific analyses. The "city" variable combines borough and neighborhood info, making it unusable. Instead, for analysis, use the "neighborhood" variable for micro-neighborhood-level analysis and the "borough" variable from data cleaning.


```{r}
# Count the observations for each ZIP code
zip_code_counts <- final_data %>%
  group_by(zipcode, neighborhood, borough, city) %>%
  summarise(Observation_Count = n())

# Filter for ZIP codes with less than 60 observations
zip_count_60 <- zip_code_counts %>%
  filter(Observation_Count < 60) %>%
  arrange(desc(Observation_Count))

knitr::kable(zip_count_60)
```

As shown in the table above there are `r nrow(zip_count_60)` zipcodes with fewer than 60 observations for change of address. There are various potential factors that are likely contributing to these specific zip codes having less observations. For example, many of these zipcodes are within the midtown and lower Manhattan neighborhoods, which are characterized by large business/commercial areas. Zipcodes in these business-centric areas may be predominantly associated with commercial addresses and PO boxes, resulting in fewer residential address changes and contributing to missing neighborhood values. Additionally, many of these zipcodes are within residential areas of Queens and Brooklyn where property ownership is more likely, thus these zipcodes may have fewer COAs.
 
 
## Section 2 - Exploratory Data Analysis and Visualization

```{r}
# Group by borough and year, and calculate the average net_change
average_net_change_table <- final_data %>%
  group_by(borough, year) %>%
  summarize(Average_Net_Change = mean(net_change, na.rm = TRUE)) %>% 
  arrange(borough, year)

average_net_change_table <- average_net_change_table %>%
  rename(`Borough` = borough, `Year` = year, `Average Net Change` = Average_Net_Change)

knitr::kable(average_net_change_table, caption = "This table shows the average net change in addresses by borough and year. The 'Average Net Change' values represent the average address changes for each combination of borough and year. Note that missing values are excluded from the calculations") 
```

Notable trends include a sharp decline in net change values for all boroughs in 2020, except Staten Island, where this occurred in 2021. In 2020, Manhattan had the most significant decline, followed by Brooklyn, the Bronx, and Queens. However, Manhattan also displayed the most robust recovery in 2021. 

```{r}
lowest_net_change <- final_data %>%
  arrange(net_change) %>%
  head(5) %>%
  select(zipcode, neighborhood, year, month, net_change)

knitr::kable(lowest_net_change, caption = "Lowest Net Change in COA Requests")
```

```{r}
highest_net_change_pre_2020 <- final_data %>%
  filter(year < 2020) %>%
  arrange(desc(net_change)) %>%
  head(5) %>%
  select(zipcode, neighborhood, year, month, net_change)

knitr::kable(highest_net_change_pre_2020, caption = "Highest Net Change in COA Requests Pre-2020")
```

```{r}

# Step 1: Filter the data by borough
manhattan_data <- final_data %>% filter(borough == "Manhattan")
brooklyn_data <- final_data %>% filter(borough == "Brooklyn")
bronx_data <- final_data %>% filter(borough == "Bronx")
queens_data <- final_data %>% filter(borough == "Queens")
staten_island_data <- final_data %>% filter(borough == "Staten Island")


# Step 2: Calculate the average net_change values across ZIP codes within neighborhoods for each month and year
manhattan_avg_net_change <- manhattan_data %>%
  group_by(year, month, neighborhood) %>%
  summarize(avg_net_change = mean(net_change)) %>%
  ungroup()

brooklyn_avg_net_change <- brooklyn_data %>%
  group_by(year, month, neighborhood) %>%
  summarize(avg_net_change = mean(net_change)) %>%
  ungroup()

bronx_avg_net_change <- bronx_data %>%
  group_by(year, month, neighborhood) %>%
  summarize(avg_net_change = mean(net_change)) %>%
  ungroup()

queens_avg_net_change <- queens_data %>%
  group_by(year, month, neighborhood) %>%
  summarize(avg_net_change = mean(net_change)) %>%
  ungroup()

staten_island_avg_net_change <- staten_island_data %>%
  group_by(year, month, neighborhood) %>%
  summarize(avg_net_change = mean(net_change)) %>%
  ungroup()


# Step 3: Create separate plots for each borough

manhattan_plot <- manhattan_avg_net_change %>%
  ggplot(aes(x = month, y = avg_net_change, group = neighborhood, color = neighborhood)) +
  geom_line() +
  geom_point() +
  facet_wrap(~year, nrow = 1) +  # Facet by year
  labs(title = "Manhattan", x = "Month", y = "Average Net Change") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

brooklyn_plot <- brooklyn_avg_net_change %>%
  ggplot(aes(x = month, y = avg_net_change, group = neighborhood, color = neighborhood)) +
  geom_line() +
  geom_point() +
  facet_wrap(~year, nrow = 1) +  # Facet by year
  labs(title = "Brooklyn", x = "Month", y = "Average Net Change") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

bronx_plot <- bronx_avg_net_change %>%
  ggplot(aes(x = month, y = avg_net_change, group = neighborhood, color = neighborhood)) +
  geom_line() +
  geom_point() +
  facet_wrap(~year, nrow = 1) +  # Facet by year
  labs(title = "Bronx", x = "Month", y = "Average Net Change") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

queens_plot <- queens_avg_net_change %>%
  ggplot(aes(x = month, y = avg_net_change, group = neighborhood, color = neighborhood)) +
  geom_line() +
  geom_point() +
  facet_wrap(~year, nrow = 1) +  # Facet by year
  labs(title = "Queens", x = "Month", y = "Average Net Change") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

staten_island_plot <- staten_island_avg_net_change %>%
  ggplot(aes(x = month, y = avg_net_change, group = neighborhood, color = neighborhood)) +
  geom_line() +
  geom_point() +
  facet_wrap(~year, nrow = 1) +  # Facet by year
  labs(title = "Staten Island", x = "Month", y = "Average Net Change") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# Step 4: Combine the plots using patchwork
library(patchwork)
all_boroughs_plot <- manhattan_plot / brooklyn_plot / bronx_plot / queens_plot / staten_island_plot

all_boroughs_plot <- all_boroughs_plot +
  plot_annotation(title = "Neighborhood-Level Average Net Change by Borough, 2018-2022")

ggsave("results/all_boroughs_plot.png", all_boroughs_plot, width = 16, height = 26, units = "in", bg = "white")

```



This dataset is limited in providing a comprehensive understanding of zipcode-level population changes because it does not account for variations in the number of individuals residing at each address which would be necessary for population change estimates. 




